Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-12-20 WSSSPE3/report/WSSSPE3_review.tex
some general cleanup/consistency

\item \textbf{Dan Gunter, Adam Arkin, Rick Stevens, Robert Cottingham, and
Sergei Maslov: \textit{Challenges of a Sustainable Software Platform for
Predictive Biology: Lessons Learned on the KBase Project}} In this talk, Dan
Gunter presented experiences and lessons learned as part of KBase, an open
software and data platform for addressing the grand challenge of systems
biology: predicting and designing biological function. KBase is a unified system
that integrates data and analytical tools for comparative functional genomics of
microbes, plants, and their communities. However, it is also a collaborative
environment for sharing methods and results, and placing those results in the
context of knowledge in the field. Being a large, multi-institutional project,
one of the big challenges is to agree on standard to enable a single,
maintainable system. Working in isolation does not work (anymore) within this
field, and the community in the field also didn't have standards for software
engineering. This is contrasted to computer science research, where these
approaches shorten design cycles, leading to more time for highly rewarded
activities like publishing, performance studies, graduating students, or
protecting ideas before publication. Instead of this more
traditional approach, KBase uses a variation on the ``Scrum'' methodology.
After picking projects and team members, 4-5 teams work on projects about two
weeks before a 1-week evaluation by the executive committee, and after which new
teams and projects might be chosen to restart the ``agile'' development cycle.
This process is intentionally open and documented.

>>>>

\item \textbf{Dan Gunter, Adam Arkin, Rick Stevens, Robert Cottingham, and
Sergei Maslov: \textit{Challenges of a Sustainable Software Platform for
Predictive Biology: Lessons Learned on the KBase Project}} In this talk, Dan
Gunter presented experiences and lessons learned as part of KBase, an open
software and data platform for addressing the grand challenge of systems
biology: predicting and designing biological function. KBase is a unified system
that integrates data and analytical tools for comparative functional genomics of
microbes, plants, and their communities. However, it is also a collaborative
environment for sharing methods and results, and placing those results in the
context of knowledge in the field. Being a large, multi-institutional project,
one of the big challenges is to agree on standard to enable a single,
maintainable system. Working in isolation does not work (anymore) within this
field, and the community in the field also didn't have standards for software
engineering. This is contrasted to computer science research, where these
approaches shorten design cycles, leading to more time for highly rewarded
activities like publishing, performance studies, graduating students, or
protecting ideas before publication. Instead of this more
traditional approach, KBase uses a variation on the ``Scrum'' methodology.
After picking projects and team members, four to five teams work on projects for about two
weeks before a one-week evaluation by the executive committee, and after which new
teams and projects might be chosen to restart the ``agile'' development cycle.
This process is intentionally open and documented.

### 2015-12-20 WSSSPE3/report/wg_main_best_practices.tex
accepting change and replicating to appendix
some general cleanup/consistency

Core questions that will need to be explored are in reliability (reproducibility), usability, extensibility, knowledge management and continuity (transitions between people). Answers to these will guide us on how a software tool becomes part of the core workflow of well identified users (stakeholders) relating to tool success and hence sustainability. 
\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability -- economic, technical, environmental and
obsolescence. 
\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}

>>>>

Core questions that will need to be explored are in reliability (reproducibility), usability, extensibility, knowledge management and continuity (transitions between people). Answers to these will guide us on how a software tool becomes part of the core workflow of well identified users (stakeholders) relating to tool success and hence sustainability. 
%\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology);
\item Dimensions of sustainability -- economic, technical, environmental and
obsolescence. 
%\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}
